---
title: "House Price in Vancouver"
author: "Jin Seo Jo"
date: "28/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(fpp2)
library(fpp3)
library(ggfortify)
library(forecast)
library(ggthemes) # for theme_fivethirtyeight
library(gridExtra) # for grid.arrange
library(tseries) # for adf.test
```

## Data used in conventional statistics
Conventional statistical analysis assumes data is identically and independently distributed (IID). In other words, observation have a constant mean (variance), and are independent.
```{r}
house_data <- ts(read.csv("House_Price_Index.csv"), frequency = 12, start = c(1990, 7))
ts.dat <- fortify(house_data)
len <- length(ts.dat$Index)
set.seed(5678)
X <- rnorm(len)
benchmark.ts  <- ts(X, frequency = 12, end = c(2020, 4))
autoplot(benchmark.ts) +
  ggtitle("IID Gaussian Time Series") +
  labs(caption = "simulated standard normal random variables") +
  theme_minimal()
```

## Stylized fact of time series
Time series data usually exhibit gradual (or secular) change in mean, seasonal/cycical change in mean, and observations are correlated.

### Trend
Victoria (black) and Vancouver (navy) (monthly) House Price Index from July 1990 to April 2020.
```{r}
ggplot(data = ts.dat) +
  geom_line(aes(x = Index, y = Victoria)) +
  geom_line(aes(x = Index, y = Vancouver), col = "navy") +
  theme_fivethirtyeight() +
  ggtitle("House Price Index") +
  labs(caption = c("black: Victoria, navy: Vancouver"))
```

### Seasonal/Cyclic
Victoria (black) and Vancouver (navy) (monthly) Sales Pair Counts from July 1990 to April 2020.
```{r}
ggplot(data = ts.dat) +
  geom_line(aes(x = Index, y = Victoria_count)) +
  geom_line(aes(x = Index, y = Vancouver_count), col = "navy") +
  theme_fivethirtyeight() +
  ggtitle("BC sales pair counts")
```
```{r}
vc_count <- ts(ts.dat$Vancouver_count, frequency = 12, end = c(2020, 4))
vic_count <- ts(ts.dat$Victoria_count, frequency = 12, end = c(2020, 4))

ggsubseriesplot(vc_count) + 
  theme_fivethirtyeight() +
  ggtitle("Vancouver (sales pair counts)")
```

Simulated white noise time series again
```{r}
ggsubseriesplot(benchmark.ts) +
  theme_fivethirtyeight() +
  ggtitle("IID Gaussian time series")
```

The vic_elec data contains half-hourly electricity demand for the state of Victoria, Australia
```{r}
vic_elec %>% gg_season(Demand, period = "week") +
  theme_fivethirtyeight() +
  theme(legend.position = "none")
```

### Temporal correlation/ serial correlation/ autocorrelation
Simulated IID N(0,1) over two consecutive observations
```{r}
df0 <- data.frame(X0 = X[-1], X1 = X[-len])

ggplot(data = df0) +
  geom_point(aes(x = X1, y = X0)) +
  ggtitle("Simulated IID N(0,1)")
```

Sales pair counts over two consecutive months
```{r}
df <- data.frame(vc_count0 = vc_count[-1], vc_count1 = vc_count[-len])
df1 <- data.frame(vic_count0 = vic_count[-1], vic_count1 = vic_count[-len])
g1 <- ggplot(data = df) +
  geom_point(aes(x = vc_count1, y = vc_count0)) +
  ggtitle("Vancouver")
g2 <- ggplot(data = df1) +
  geom_point(aes(x = vic_count1, y = vic_count0)) +
  ggtitle("Victoria")
grid.arrange(g1, g2, ncol = 2)
```

Autocorrelation functions
```{r}
ggAcf(vc_count) +
  theme_fivethirtyeight() +
  ggtitle("Vancouver")
```
```{r}
ggAcf(vic_count) +
  theme_fivethirtyeight() +
  ggtitle("Victoria")
```
```{r}
ggAcf(benchmark.ts) +
  theme_fivethirtyeight() +
  ggtitle("Simultaed IID N(0,1)")
```
```{r}
vic_elec %>% gg_season(Demand, period = "week") + 
  theme_fivethirtyeight() +
  theme(legend.position = "none")
```

### Unstable (unstationary)
```{r}
adf.test(benchmark.ts)
```

## Time Series Decomposition
```{r}
vc_count %>% stl(t.window = 13, s.window = "periodic", robust = TRUE) %>% 
  autoplot() +
  ggtitle("Sales pair counts (Vancouver)") +
  theme_fivethirtyeight()
```

## Forecasting
```{r}
fit1 <- stlf(vc_count, method = "rwdrift")
fit2 <- stlf(vc_count, method = "arima")
fit3 <- stlf(vc_count, method = "ets")
fit4 <- stlf(vc_count, method = "naive")

l1 <- autoplot(fit1) + ggtitle("random walk") + theme_fivethirtyeight()
l2 <- autoplot(fit2) + ggtitle("arima") + theme_fivethirtyeight()
l3 <- autoplot(fit3) + ggtitle("ets") + theme_fivethirtyeight()
l4 <- autoplot(fit4) + ggtitle("naive") + theme_fivethirtyeight()

grid.arrange(l1, l2, l3, l4, ncol = 2)
```

