---
title: "Forecasting: Principles and Practice"
author: "Jin Seo Jo"
date: "29/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(fpp2)
library(ggfortify)
library(forecast)
```
```{r, include=FALSE}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Examples of time series data
### Annual data
#### Electricity sales
Annual **electricity sales** to residential custormers in South Australia from 1989 to 2008. Electricity (used for hot water has been excluded)
```{r}
elecsales
```

```{r}
autoplot(elecsales) +
  ggtitle("Figure 1. Annual electricity sales") +
  ylab("Price") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

#### Canadian Lynx trappings
Annual **numbers of lynx trappings** for 1821 - 1934 in Canada.
```{r}
autoplot(lynx) +
  ggtitle("Figure 2. Annual Canadian Lynx Trappings") +
  xlab("Year") +
  ylab("Number trapped") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

### Quarterly Data
#### Australian GDP per capita, 1971:Q3 - 1998:Q1
```{r}
head(ausgdp, 14)
```

```{r}
# start(ausgdp) # 1971 3
# end(ausgdp) # 1998 1
# frequency(ausgdp) # 4
autoplot(ausgdp) +
  ggtitle("Figure 3. Quarterly Australian GDP per capita") +
  ylab("GDP per capital") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

#### Beer production
Total quarterly beer production in Australia (in megalitres) from 1956:Q1 to 2010:Q2
```{r}
beer <- window(ausbeer, start = 1992)

autoplot(beer) +
  ggtitle("Figure 4. Quarterly beer production (Time series plot)") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

```{r}
ggsubseriesplot(beer) +
  ggtitle("Figure 5. Quarterly beer production (Subseries plot)") +
  theme(plot.title =  element_text(size = 8, face = "bold"))
```

```{r}
ggseasonplot(beer, polar = TRUE) +
  ggtitle("Figure 6. Quarterly beer production (Season plot)") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

```{r}
# Alternative option
ggseasonplot(beer, year.labels = TRUE) +
  ggtitle("Figure 7. Quarterly beer production (Season plot)") +
  theme(legend.text = element_text(size = 8, face = "bold"))
```

### Monthly data
#### Air Passengers
Monthly totals of international airline passengers from 1949 to 1960.
```{r}
data("AirPassengers")

# start(AirPassengers)
# end(AirPassengers)
# frequency(AirPassengers)

p <- autoplot(AirPassengers) +
  ggtitle("Figure 8. Monthly totals of international airline passengers") +
  ylab("Number of passengers (1000's)") +
  theme(plot.title = element_text(size = 8, face = "bold"))

p
```

```{r}
ggsubseriesplot(AirPassengers) +
  ylab("1000's") +
  ggtitle("Figure 9. Monthly airline passengers") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

```{r}
ggseasonplot(AirPassengers, polar = TRUE) +
  ggtitle("Figure 10. Monthly airline passengers (Season plot 1)") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

```{r}
ggseasonplot(AirPassengers, year.labels = TRUE, year.labels.left = TRUE) +
  ggtitle("Figure 11. Monthly airline passengers (Season plot 2)") +
  ylab("Number of passengers (1000's)") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

#### Road causalities in Great Britain 1969-84
UKDriverDeaths is a time seires giving the monthly totals of car drivers in Great Britain killed or seriously injured Jan 1969 to Dec 1984. Compulsory wearing of seat belts was introduced on 31 Jan 1983. 
```{r}
data("UKDriverDeaths")

autoplot(UKDriverDeaths) +
  ggtitle("Figure 12. Road causalities in Great Britain") +
  ylab("Number of deaths") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

```{r}
ggsubseriesplot(UKDriverDeaths) +
  ggtitle("Figure 13. Road causalities in Great Britain (Subseries plot)") +
  ylab("Number of deaths") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

```{r}
ggseasonplot(UKDriverDeaths, polar = TRUE) +
  ggtitle("Figure 14. Road causalities in Great Britain (Season plot)") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

### Other frequency
US treasury bill contracts on the Chicago market for 100 consecutive trading days in 1981.
```{r}
autoplot(ustreas) +
  ggtitle("Figure 15. US Treasury Bill Contracts") +
  xlab("Day") +
  ylab("Price") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

## Patterns in time series data
Three patterns are observed in time series data:  
1. Trend pattern exists when there is a long-term increase or decrease in the data.  
2. Seasonal pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).  
3. Cyclical pattern exists when data exhibit rises and falls that are not of fixed period (duration usually of at least 2 years)

## Measuring temporal dependence
### Common tools
```{r}
beer <- window(ausbeer, start = 1992)

p00 <- autoplot(beer) +
  ggtitle("Beer production") +
  theme(plot.title = element_text(size = 8, face = "bold"))

p01 <- ggsubseriesplot(beer) +
  ggtitle("Beer production (Subseries plot)") +
  theme(plot.title = element_text(size = 8, face = "bold"))

multiplot(p00, p01, cols = 2)
```

1. **Lagged scatterplot**: Each graph shows $y_{t}$ plotted against $y_{y-k}$ for different values of k.
```{r}
forecast::gglagplot(beer) +
  ggtitle("Figure 16, Beer production (Lagged plot)") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

2. **Autocorrelation plots (aka correlogram)**: The autocorrelations are the correlations associated with these scatterplots. Mathematically, the autocorrelation function at lag k may be defined as
$$p(k) = \frac{cov(y_{t},y_{t-k})}{\sqrt{var(y_{t})var(y_{t-k})}}$$
```{r}
ggAcf(beer) +
  ggtitle("Figure 17. Beer production (ACF)") +
  theme(plot.title = element_text(size = 8, face = "bold"))
```

### Autocorrelation plots for a white noise process
A stochastic process $\{y_{t}\}$ is called a white noise process if $y_{t}$ satisfies  
1. $E(y_{t}) = 0 \; \forall t$  
2. $var(y_{t}) = cov(y_{t}, y_{t}) = {\sigma^2}$ is a constant.  
3. $\gamma(t,s) = cov(y_{t}, y_{s}) = 0 \; \forall t \neq s$

```{r}
set.seed(1234)

y <- ts(rnorm(100))

p1 <- autoplot(y) +
  ggtitle("Time series plot of white noise") +
  theme(plot.title = element_text(size = 8, face = "bold"))

p2 <- ggAcf(y, lag.max = 20) +
  ggtitle("White noise (ACF)") +
  theme(plot.title = element_text(size = 8, face = "bold"))

multiplot(p1, p2, cols = 2)
```

### Autocorrelation plos for data with trend and seasonality
```{r}
elec <- window(elec, start = 1980)

p3 <- autoplot(elec) +
  ggtitle("Electricity production") +
  theme(plot.title = element_text(size = 8, face = "bold"))

p4 <- ggAcf(elec, lag.max = 48) +
  ggtitle("Electricity production (ACF)") +
  theme(plot.title = element_text(size = 8, face = "bold"))

multiplot(p3, p4, cols = 2)
```

**Remark:**  
1. When data have a trend, the autocorrelations for small lags tend to be large and positive.  
2. When data are seasonal, the autocorrelations will be larger at the seasonal lags (i.e., at multiples of the seasonal frequency)  
3. When data are trended and seasonal, you see a combination of these effects. 
