---
title: "Modeling Irregular Component - ARMA model"
author: "Jin Seo Jo"
date: "06/10/2020"
output: html_output
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, message=FALSE}
library(fpp3)
library(ggplot2)
library(tidyverse)
library(ggfortify)
library(tidyquant)
library(forecast)
library(dygraphs)

library(corrplot)
library(gridExtra)
library(knitr)
```

## ARMA model identificataion
### Mathematics behind ARMA model identification
Consider
$$X_{t} - \phi_{1}X_{t-1} - ... - \phi_{p}X_{t-p} = a_{t} + \theta_{1}a_{t-1} + ... + \theta_{q}a_{t-q}, \\ a_{t} \sim NID(0, \sigma_{a}^2)$$
where $NID(0, \sigma_{a}^2)$ denotes IID normal random variables with mean zero and variable $\sigma_{a}^2$.  
Note that in this document we assume $a_{t} \sim NID(0, \sigma_{a}^2)$

### Use R to check your answers
Calculate theorethical ACF and PACF of an ARMA model in R:  
`ARMAacf(ar = numeric(), ma = numeric(), lag.max = r, pacf = FALSE)`

#### Theoretical ACF and PACF of AR(1) process
```{r}
ar.acf <- zapsmall(ARMAacf(ar = 0.5, lag.max = 5))[-1]
ar.pacf <- zapsmall(ARMAacf(ar = 0.5, lag.max = 5, pacf = TRUE))
ar.df <- data.frame(ACF = ar.acf, PACF = ar.pacf, lag = 1:5)

p1<- ggplot(data = ar.df, aes (x = lag, y = ACF)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = ACF), vjust = 2, size =2, color = "white") +
  theme_minimal() +
  ggtitle("ACF")

p2<- ggplot(data = ar.df, aes (x = lag, y = PACF)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = PACF), vjust = 2, size =2, color = "white") +
  theme_minimal() +
  ggtitle("PACF")

grid.arrange(p1,p2, nrow = 1)
```

#### Theoretical ACF and PACF of MA(1) process
```{r}
ma.acf <- round(zapsmall(ARMAacf(ma = 0.5, lag.max = 5))[-1], 2)
ma.pacf <- round(zapsmall(ARMAacf(ma = 0.5, lag.max = 5, pacf = TRUE)), 2)
ma.df <- data.frame(ACF = ma.acf, PACF = ma.pacf, lag = 1:5)

p3 <- ggplot(data = ma.df, aes(x = lag, y = ACF)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = ACF), vjust = 2, size = 2, color = "white") +
  theme_minimal() +
  ggtitle("ACF")

p4 <- ggplot(data = ma.df, aes(x = lag, y = PACF)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = PACF), vjust = 2, size = 2, color = "white") +
  theme_minimal() +
  ggtitle("PACF")

grid.arrange(p3, p4, nrow = 1)
```

Autocorrelation matrix as Toeplitz matrix
```{r}
m <- toeplitz(ARMAacf(0.8, lag.max = 7))
rownames(m) = 0:7
colnames(m) = 0:7
kable(m, digits = 2, caption = "Autocorrelation matrix", format = "markdown")
```

```{r}
corrplot(m)
```

### Rule of Thumb
||ACF|PACF|
|:--:|:--:|:--:|
|AR(p)|Tails off or damped sine wave|Cut off after lag p|
|MA(q)|Cut off after lag q|Tails off or damped sine wave|
|ARMA(p,q)|Tails off after lag(q-p)|Tails off after lag(p-q)|

```{r}
set.seed(5678)
sim.ar1 <- arima.sim(n = 100, list(ar = 0.5))
sim.ma1 <- arima.sim(n = 100, list(ma = 0.5))

par(mfrow = c(2,2), cex = 0.5)
acf(sim.ar1, main = "ACF & AR (1)", col = 2);pacf(sim.ar1, main = "ACF & AR (1)", col = 2)
acf(sim.ma1, main = "ACF & MA (1)", col = 2);pacf(sim.ma1, main = "ACF & MA (1)", col = 2)
```

## Checking stationary of ARMA models
### Stationary of ARMA models
**How can we tell if a given ARMA model is stationary?**  
Show that a given ARMA model satisfying the definition of weak stationary.  

**A general approach to test stationary of an ARMA model**
Consider an ARMA(p,q) model
$$\phi(B)X_{t} = \theta(B)a_{t}$$
where
$$\phi(B) = 1 - \phi_{1}B - \phi_{2}B^2 - ... - \phi_{p}X_{t-p}, \\ \theta(B) = 1 + \theta_{1}B + \theta_{2}B^2 + ... + \theta_{q}B^q$$
and $B^kX_{t} = X_{t-k}$ is the backward shift operator.  

**If an ARMA process defined in the equation is stationary, all roots in $\phi(B)$  must lie outside a unit circle.**

### Simulated examples
Consider the following simulated ARMA processes
$$ X_{t} - 0.5X_{t-1} = a_{t}, \; a_{t} \sim NID(0,1) \\
X_{t} = a_t + 0.7a_{t-1}, \; a_{t} \sim NID(0,1) \\
X_{t} - 0.5X_{t-1} = a_t + 0.7a_{t-1}, \; a_{t} \sim NID(0,1)$$

```{r}
# Simulated time series
set.seed(1234)
ar1 <- arima.sim(list(order = c(1,0,0), ar = 0.5), n = 200)
ma1 <- arima.sim(list(order = c(0,0,1), ma = 0.7), n = 200)
arma11 <- arima.sim(list(order = c(1,0,1), ar = 0.5, ma = 0.7), n = 200)

autoplot(cbind(ar1, ma1, arma11)) +
  ggtitle("Simulated stationary time series") +
  theme_minimal() +
  ylab("") +
  theme(legend.position = "bottom")
```

AR(1) model
```{r}
mod.ar1 <- arima(ar1, c(1,0,0))
coef(mod.ar1) %>% kable()
plot(mod.ar1)
```

MA(1) model
```{r}
mod.ma1 <- arima(ma1, c(0,0,1))
coef(mod.ma1) %>% kable()
plot(mod.ma1)
```

ARMA(1,1) model
```{r}
mod.arma11 <- arima(arma11, c(1,0,1))
coef(mod.arma11) %>% kable()
plot(mod.arma11)
```

## Causal and invertible representation of an ARMA model
### Definition
- An ARMA model is a causal process if it can be expressed as a stationary $MA(\infty)$ process
$$X_{t} = \sum_{i=0}^{\infty} \psi_{i}a_{t-i}$$
- An ARMA model is an invertible process if it can be expressed as a stationary $AR(\infty)$ process
$$X_{t} - \sum_{i=0}^{\infty} \pi_{i}X_{t-i} = a_{t}$$
- We can use R to calculate $\psi$ coefficients.

### Nonstationary ARMA model
Consider the following ARMA model
$$ X_{t} - 1.2X_{t-1} = a_{t} + 0.5a_{t-1}$$
```{r}
library(FitARMA)

psi <- ImpulseCoefficientsARMA(phi = 1.2, theta = 0.5, lag.max = 10) %>% 
  round(digits = 3)

psi.df <- data.frame(lag = 0:(length(psi)-1), psi = psi)

ggplot(data = psi.df, aes(x = lag, y = psi)) +
  geom_bar(stat = "identity", alpha = 0.3) +
  geom_text(aes(label = psi), vjust = 1, size = 2, color = "black") +
  theme_minimal() +
  ggtitle("Psi coefficients")
```

